\textcolor{blue}{A gradient-based optimization method for motion planning}$^{1}$

### 2016-01-25 answers-to-reviewers.tex

\newcommand{\todo}{\textcolor{red}{TODO}}

%\bibliographystyle{unsrt}
%\bibliographystyle{IEEEtran}
\bibliographystyle{tADR}
\bibliography{paper}
%\bibliography{IEEEabrv,paper}

\begin{document}

\title{A simple path-optimization method for motion planning}
\author{Myl\`{e}ne Campana, Florent Lamiraux and Jean-Paul Laumond}
\date{Answer to reviewers}
\maketitle

@@

\subsection{Comment 2}
\hrulefill\\

Introduction: 
It would be very nice with a discussion of what optimal means. Optimal can be with respect to path length (the classical criteria also used in this paper), but could just as well be optimal with respect to execution time, where the number of via points and the clearance (allowing for larger blends) are important factors. Optimal could also depend on some task specific objectives.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

It is true that other optimization criteria can be different of the path length. 
Thus, relying on the reviewer comment, we add the mention of others. Concerning 
the choice of the path length, it allows us to obtain a quadratic cost, better 
conditioned for optimization techniques.

Finally, we have developed the meaning of optimality, and we refer to a 
recent paper ACM-Com by one of the author.

@@

\subsection{Comment 4}
\hrulefill\\

Related work: 
The work in [11] is also described in an IJRR paper of the same authors. This paper could be cited as well.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We thank the reviewer for the reference, [11] has been changed to cite the journal paper instead of the conference.

@@

\subsection{Comment 5}
\hrulefill\\

Section 3:
The paper starts describing the details of the algorithm and reference to algorithm 2 placed several pages later. It would make the paper more readable if the overall algorithm was presented earlier making the reader aware of where the components would be used before describing them. This would also mean that the reader does not have to flip back and forth all the time.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We understand that starting to define the algorithm's functions and reference 
to it too early is confusing.
We prefer not to introduce the algorithm earlier because its structure ensue 
from the problem's construction.
Instead, we prefer to remove the reference to the algorithm. We also precise at 
the beginning of the part 3 that some of the problem steps correspond to 
functions that will be called in the algorithm, given at the end. Thus we do not need to redefine completely the functions in the algorithm section.

@@

\subsection{Comment 7}
\hrulefill\\

Results:
The authors only compares their approach with random shortcut and partial random shortcut, but omit the standard path pruning technique also described in [11]. This technique works excellent for the special case in Figure 6.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

It is true that in the case of Figure 6, Prune is complementary to random shortcut. But this method, used alone, has also its counter-examples, as the one presented in Figure 1 (and the PR2-crossing-arms Figure 9).

@@

\subsection{Comment 8}
\hrulefill\\

In the paper the authors state that it can be hard to select the stopping criteria of the random short cutting algorithms, for which the authors have selected five iterations with non-improvement. Looking at the time it takes before the optimization algorithms terminates (Table 2) it appears that this criterion is too harsh as the PRS and RS in many cases terminates significantly before the GB based method. To show if their method is indeed more efficient they should try to let the RS and PRS run for the same amount of time as the GB method. It would also be nice if Table 2 would include the number of iterations and potentially selected plots showing how the path length decreases as a function of time/iterations.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We admit that the termination conditions of the algorithms were not fair. Instead, as suggested reviewers 1 and 3, we choose to allow the same amount of optimization duration for the three optimizers. Since the Gradient-Based optimizer is deterministic, its duration has been chosen.

To provide information about the ``path length decreases as a function of time'', we have added Figure 13 some graphs comparing the three optimizers.

@@

\subsection{Comment 9}
\hrulefill\\

Related to my previous comment about path pruning, then I would generally recommend people to use a combination of the deterministic path pruning followed by random short cut algorithm, as this in my experience performs better (in terms of how much it can optimize in a given time) then using just short cutting.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We thank the reviewer for the suggestion of using Prune as a preliminary optimization step for each optimizer. Regarding to our algorithm, Prune has pros and cons: its removes waypoints i.e. optimization variables so the computation time reduces. However, less waypoints results in less possible actions to blend the path. In fact, we noted that we obtained shorter paths without the pruning stage.

\todo : voir si ajout d'une partie "comparaison avec et sans prune" .... globalement, prune ne change pas grand chose a la comparaison avec les autres optim, si ce n'est qu'on a en moyenne des chemins initiaux similaires (peu de waypoints)

@@

It looks that the conference version of this paper [1] is not cited in the main text though it appears in the Reference section.  Please add the citation in an appropriate place.

[1] Campana M, Lamiraux F, Laumond JP. A simple path optimization method for 
motion planning. https://hal.archives-ouvertes.fr/hal-01220048. 2015.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We thank the reviewer for the comment. We removed the reference from the paper.

@@

1) Notation and clarity of the presentation:

When introducing the Jacobian of f, at no point has f been introduced as the constraint function. I think in general the problem should be introduced as an optimization problem, what is the objective, what are the constraints, what simplifications will be made.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We thank the reviewer for the remark. We have improved the notations of the problem, avoiding to introduce unreferenced objects at the beginning.

\todo : adapter / etoffer en fonction de ce que Florent a re-modifie

@@

\subsection{Comment 2}
\hrulefill\\

In that regard it would be nice to comment the Newton method, for example there exists many methods to select alpha. Having a constant alpha the most naïve approach.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We think that the reviewer is referring to line-search algorithms to 
evaluate alpha. However these algorithms are not relevant for a QP problem since 
the cost minimum is known. We just intend to process \textit{small} steps toward it, and to note collision-constraints. We admit that the alpha value could be chosen in a smarter way, e.g. due to geometrical considerations, but we mentioned it in the conclusion as a future work.

\todo : verifier pertinence

@@

\subsection{Comment 3}
\hrulefill\\

I wonder if the discussion concerning the configuration space, which is well done by the way, is of importance here, it is a perquisite in the robotics community.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We want to highlight the fact that we optimize paths on the robot 
configuration space in a proper mathematical way, specially regarding to the $SO3$ rotations. Thus we prefer to briefly define notations and operators.

@@

\subsection{Comment 4}
\hrulefill\\

The iteration index notation and the time index use the same letter (i), making it difficult to follow. The p variable is sometimes called back with mention of what it is.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We apologize for the unclear/unreferenced notations and have improved it.

@@

\subsection{Comment 5}
\hrulefill\\

The kernel discussion, and velocity discussion are a bit breve and unclear for the reader.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

The authors agree with the reviewer that this part should be modified as it is unclear. We add the Figure 5 to help understanding the interest of the 3.6 part.

\todo : adapter en fonction de ce que Florent a re-modifie

@@

The experimental section leaves the impression that the comparisons with other methods are either giving the advantage to the state-of-the-art or, simply unfair.

Tuning the shortcut method by changing the distribution of samples along the path drastically changes the converging behavior, so choosing '5 non improving iterations' as the stopping criterion is very arbitrary. And it is the case in the results that the GB method beats the other methods using 1000 times the budget allocated to the other methods... I think reducing the comparison to one number in this case does not allow the reader to get a fair estimate of the convergence rate. A common criterion to be used is simply a max iteration count, which would already be fairer, but in order to not have these criteria influence the comparison, why not plot the averaged convergence curves?

\rule{\linewidth}{.1pt}
\textbf{Answer}:

The unfair comparison, also raised by the reviewer 1, has been improved: we allocate the Gradient-Based duration as the computation times for the random optimizers. Thus we avoid arbitrary termination condition.

The drawback of this method is that we ignore if the computation time of our method is too important. Thus, as suggested, we have represented the convergence curves for some benchmarks, to see if RS and PRS are not converging far rapidly than GB.

@@

\subsection{Comment 7}
\hrulefill\\

Otherwise it would be interesting to see the alpha parameter being discussed at that point in the paper (with experimental results to back up the discussion).

\rule{\linewidth}{.1pt}
\textbf{Answer}:

As suggested, we have added at the end of the results some graphs for different values of $\alpha_{init}$.

### 2016-01-25 paper.tex
Rewrite sections 3.5, 3.6.

% Math shortcuts :
\newcommand\real{\mathbb{R}}
\newcommand\p{\mathbf{p}}
\newcommand\gT{\tilde{g}^T}
\newcommand\g{\tilde{g}}
\newcommand\CS{\mathcal{C}}
\newcommand\dimCS{n_\mathbf{C}}
\newcommand\body{{\cal B}}
\newcommand\Sone{\mathbf{S}^1}
\newcommand\Sthree{\mathbf{S}^3}
\newcommand\conf{\mathbf{q}}
\newcommand\xx{\mathbf{x}} % \x already defined in the class
\newcommand\cost{C}
\newcommand\weight{W}
\newcommand\translation{\mathbf{t}}
%\newcommand\tcolli{t_{coll\ i}}
\newcommand\tcolli{\kappa_i}
\newcommand\po{\mathbf{P}}
\newcommand\Jf{\Phi}
\newcommand\kernel{\mbox{Ker }}
\newcommand\U{\mathbf{u}}
\newcommand\rotation{R}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\graphicspath{{images/}}

@@

\subsubsection{New constraint}

@@

\begin{figure}
	\centering
	\includegraphics[width=15.8cm]{contact-points.pdf}
	\caption{Bodies representation in collision-configuration (left) and 
	backtracked free-configuration (right). A contact point (i.e. any point in 
	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
	Collision Library~\cite{fcl}, used to detect collisions. 
%	The red $(\perp)$ 
%	hyperplan is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
%	direction which is allowed by the constraint.
	}
	\label{contact-points}
\end{figure}

@@

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame 
(Figure~\ref{contact-points} left).
\textcolor{red}{We denote by
  \begin{itemize}
  \item $\po_1$ (resp. $\po_2$) the point belonging to body $\body_1$ (resp. $\body_2$) that coincides with $\po_c$ in configuration $\xx_{i+1}(\tcolli)$,
  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame in configuration $\conf$,
  \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
  \end{itemize}
  The position of $\po_1$ (resp. $\po_2$) in the global reference frame depends on the configuration $\conf$ of the robot. We therefore denote $\po_1(\conf)$ (resp. $\po_2(\conf)$) this position.
We define $\U$ as the unit vector attached to $\body_1$, linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$ (Figure~\ref{contact-points} right):
$$
\U (\conf) = \rotation_1(\conf)\rotation_1^{T}(\xx_{i}(\tcolli))\frac{\po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))}{\|\po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))\|}
$$
Note that
\begin{itemize}
\item $\U$ depends on the configuration of the robot $\conf$, since $\U$ moves with $\body_1$,
\item $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
\end{itemize}
}
\textcolor{blue}{
Let $g$ be the function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
$$
g (\conf) = \left(\po_2(\conf) - \po_1(\conf) | \mathbf{u}\right)
$$
}
%which is calculated from the last collision-free path at abscissa $\tcolli$.

Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
$$
f (\xx) = g(\xx (\tcolli))
$$
The constraint defined for any path $\xx$ by:
\begin {equation}\label{eq:new-constraint}
f(\xx) - f(\xx_{i}) = 0.
\end {equation}
aims at keeping colliding points $P_1$ and $P_2$ in configuration $\xx_{i+1} (\tcolli)$ apart in direction $\U$ defined by those points in configuration $\xx_{i} (\tcolli)$.
We linearize the constraint around $\xx_{i}$: 
$$
\frac{\partial f}{\partial \xx}(\xx_i)(\xx - \xx_i) = 0
$$
The computation of the linearized constraint is described in the next
section. Then, a line is added in the constraint Jacobian matrix $\Jf_i$:
\begin {align*}
\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{j+1}\end{array}\right)\ \mbox {with}\\
L_{j+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
\end{align*}
\textcolor{blue}{This step is denoted by the function \texttt{addCollisionConstraint}}.

@@

Thus the linearized constraint Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ 
is built by matrix 
blocks using the Jacobian $\frac{\partial f}{\partial \conf}$ expressed in 
each of the two waypoints and $\beta$. This stage \textcolor{blue}{is refered as 
to \texttt{computeCollisionConstraint}}.

\subsection{Convergence Analysis}

\begin {figure}
  \centerline {
    \def\svgwidth {.7\linewidth}
    \graphicspath{{./images/}}
    \input {images/convergence.pdf_tex}
  }
  \caption {Illustration of linearized constraints: we want to minimize the quadratic cost $\frac{1}{2}\|\xx-\xx^*\|^2$ under the non-linear constraint $f(\xx)\leq 0$. $\alpha_{init}=\frac{1}{4}$. The algorithm starts from $\xx_0$. The first iterate is $\xx_1$ which satisfies the constraint. The second iterate is $\xx_2$ that does not satisfy the constraint. The algorithm backtracks to $\xx_1$ and inserts linear constraint $L_1: \frac{\partial f}{\partial \xx} (\xx_1)(\xx - \xx_1) = 0$. $\xx_3$ is the global minimum under $L_1$. As $\xx_3$ does not satisfy the constraint, the algorithm moves to $\xx_4$ that satisfies the constraints, and then to $\xx_5$ that does not. The algorithms backtracks to $\xx_4$, inserts constraint $L_2: \frac{\partial f}{\partial \xx} (\xx_2)(\xx - \xx_2) = 0$, and returns $\xx_4$ as a solution since the dimension of the search space is 0. Notice that the same non-linear constraint may give rise to several linear constraints. The convergence of the algorithm relies on the fact that the kernel of constraint $L2$ is not contained in the kernel of the current constraints ($L_1$ only here). The convergence analysis can be roughly summarized as follows. $L2$ to be linearly dependent of $L_1$ requires that $f$ is stationary along $L_1$ at $\xx_4$. This is unlikely (but possible) since $\xx_4$ is not far from the boundary of the domain defined by $f(\xx) \leq 0$. If $L_2$ was not linearly independent from $L_1$, Algorithm~\ref{algo:solveRedundant} would keep searching new iterates between $\xx_5$ and $\xx_4$. If by any chance constraint $f$ linearized around each of those collision-free iterates was each time linearly dependent from $L_1$ the iterates would converge to the boundary of the domain defined by $f(\xx) \leq 0$. By continuous differentiability of $f$, this would mean that $f$ is stationary at the boundary. In other words, the straight line passing by $\xx_1$ and $\xx_3$ would cross the boundary tangently (as in bottom left picture). This is possible but unlikely, unless the problem has been defined as such on purpose.}
  \label {fig:convergence}
\end{figure}

@@

From the definition of $\Jf_{i}$, it is straightforward that:
\begin{equation}\label{eq:decreasing-kernel}
\kernel \Jf_{i+1} \subset \kernel \Jf_{i}
\end{equation}
\textcolor{blue}{In other words, any iteration step $\p_i$ that complies with the 
collision constraints on path $\xx_{i+1}$, necessarily sastifies $\xx_i$ constraints.}
Let us assume that:
\begin{equation}\label{eq:assumption-1}
L_{k+1}\p_i \not= 0
\end{equation}
We will elaborate later on this assumption. This means that 
$\p_{i}\notin\kernel \Jf_{i+1}$, and as $\p_{i}\in\kernel \Jf_{i}$, then:
$$
\kernel \Jf_i \not= \kernel \Jf_{i+1}
$$
From~(\ref{eq:decreasing-kernel}), we deduce that:
$$
\dim (\kernel \Jf_{i+1}) < \dim (\kernel \Jf_i)
$$
This result proves that under assumption~(\ref{eq:assumption-1}), each added constraint 
is linearly independent from the 
previous ones, and thus our algorithm terminates in a finite number of iterations.

@@

\subsubsection{Note about assumption~(\ref{eq:assumption-1})}%
\noindent
Let us consider the following trajectory:
\begin{equation}\label{eq:trajectory}
\conf (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (\tcolli), \ t \in [0,1]
\end{equation}
\textcolor{blue}{where $\xx_i$ is still collision-free and $\xx_{i+1}$ 
has its first collision detected at abcissa $\tcolli$.}
The motion $\conf(t)$ is defined by taking a constant abscissa $\tcolli$ and by 
moving from the 
path $\xx_i$ along the step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.

Thus the velocity of point $\po_2(\conf(t))$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot is given by:
$$
 \dot{\conf}(t) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_i + t\alpha_i\xx_{i+1})\p_i
$$

Therefore the following expressions
% \frac{\partial \conf}{\partial t}(0) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i})\p_i \ \text{and} \ \frac{\partial \conf}{\partial t}(1) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
$$
\dot{\conf}(0) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i})\p_i \ \ \text{and} \ \ \dot{\conf}(1) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
$$
correspond to $(\dot{\po_2}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$. The term $\dot{\conf}(1)$ being equal to zero requires either:
\begin {enumerate}
\item that $\po_1$ and $\po_2$ come to contact at zero velocity, or
\item the velocity of $\po_2$ in the reference frame of $\body_1$ is orthogonal to 
$\mathbf{u}$ at contact ($t=1$).
\end {enumerate}
The first case is very unlikely and almost never encountered in real applications, 
except if the user builds such a case on purpose.
The second case may appear and can be noted by a rank loss of $\Jf_{i+1}$.
Then we do not directly insert the new constraint. Instead, we process intermediate 
iterations until the redundancy vanishes, as explained in function 
\texttt{findNewConstraint} in Algorithm~\ref{algo:solveRedundant}, and illustrated in Figure~\ref{fig:convergence-diagram}.

After producing an intermediate iteration $\xx_{i+2}$ where $\alpha_i$ has
been halved, the obtained path is tested for collisions. If it is collision-free, the 
constraint is now linearized around it. Otherwise, new contact points are taken 
on it, and the constraint is linearized around the latest free path.

\begin{figure}
	\centering
	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
	\caption{\textcolor{blue}{Considering a trajectory 
	$\conf (t) = (\xx_{Free} + t\, \alpha \p)(\kappa)$ 
	with $\xx_{Coll} = \xx_{Free} + \alpha\p$ ($\xx_{Free}$, $\xx_{Coll}$, $\p$ 
	and $\kappa$ respectively collision-free path, in-collision path, iteration step 
	and collision-abcissa).
	The trajectory is represented in the configuration space (left) and in the 
	workspace with the colliding bodies (right).}
	Note that a case with 
	$(\dot{\po_2}(\xx_{Free})|\mathbf{u})=0$ is considered, so $\p$ already complies 
	with the constraint that should be added if no redundancy was tracked. Therefore, 
	Algorithm~\ref{algo:solveRedundant} will produce one intermediate step
	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
	$(\dot{\po_2}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
	returning the new constraint.}
	\label{fig:convergence-diagram}
\end{figure}

>>>>

% Math shortcuts :
\newcommand\real{\mathbb{R}}
\newcommand\p{\mathbf{p}}
\newcommand\gT{\tilde{g}^T}
\newcommand\g{\tilde{g}}
\newcommand\CS{\mathcal{C}}
\newcommand\dimCS{n_\mathbf{C}}
\newcommand\body{{\cal B}}
\newcommand\Sone{\mathbf{S}^1}
\newcommand\Sthree{\mathbf{S}^3}
\newcommand\conf{\mathbf{q}}
\newcommand\xx{\mathbf{x}} % \x already defined in the class
\newcommand\cost{C}
\newcommand\weight{W}
\newcommand\translation{\mathbf{t}}
%\newcommand\tcolli{t_{coll\ i}}
\newcommand\tcolli{\kappa_i}
\newcommand\po{\mathbf{P}}
\newcommand\Jf{\Phi}
\newcommand\kernel{\mbox{Ker }}
\newcommand\U{\mathbf{u}}
\newcommand\rotation{R}
\newcommand\traj{\Gamma}
\newcommand\velocity{\mathbf{v}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\graphicspath{{images/}}

@@

\subsubsection{New constraint}\label{subsubsec:new-constraint}

@@

\begin{figure}
	\centering
	\includegraphics[width=15.8cm]{contact-points.pdf}
	\caption{Bodies representation in collision-configuration (left) and 
	backtracked free-configuration (right). A contact point (i.e. any point in 
	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
	Collision Library~\cite{fcl}, used to detect collisions. Constraint $f$ defined by~(\ref{eq:f}) aims at keeping $\po_2$ in plane $\Pi$ fixed to $\body_1$.
	}
	\label{contact-points}
\end{figure}

@@

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame 
(Figure~\ref{contact-points} left).
\textcolor{red}{We denote by
  \begin{itemize}
  \item $\po_1^{loc}$ (resp. $\po_2^{loc}$) the coordinate vector of $\po_c$ in the local frame of  $\body_1$ (resp. $\body_2$),
  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame in configuration $\conf$,
  \item $M_2^1 (\conf) = M_1(\conf)^{-1} M_2(\conf)$ the position of $\body_2$ local frame in $\body_1$ local frame in configuration $\conf$,
  \item $\po_1(\conf)$ (resp. $\po_2(\conf)$) the points moving with $\body_1$ (resp. $\body_2$) of local coordinate $\po_1^{loc}$ (resp. $\po_2^{loc}$) in $\body_1$ (resp. $\body_2$) local frame.
  %% \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
  \end{itemize}
We define $\U$ as the coordinate vector of the unit vector linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$, expressed in local frame of $\body_1$ (Figure~\ref{contact-points} right):
$$
\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}\|}
$$
Note that $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
}
\textcolor{blue}{
  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
\begin{equation}\label{eq:g}
g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - P_1^{loc} | \U\right)
\end{equation}
}
%which is calculated from the last collision-free path at abscissa $\tcolli$.

Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
\begin{equation}\label{eq:f}
f (\xx) = g(\xx (\tcolli))
\end{equation}
\textcolor{red} {The constraint defined for any path $\xx$ by:
\begin {equation}\label{eq:new-constraint}
f(\xx) - f(\xx_{i}) = 0.
\end {equation}
aims at keeping point $\po_2$ in a plane fixed to $\body_1$, orthogonal to $\U$ and and passing by $\po_2$ in configuration $\xx_i(\tcolli)$ (Figure~\ref{contact-points} right).
}
We linearize the constraint around $\xx_{i}$: 
$$
\frac{\partial f}{\partial \xx}(\xx_i)(\xx - \xx_i) = 0
$$
The computation of the linearized constraint is described in the next
section. Then, a line is added in the constraint Jacobian matrix $\Jf_i$:
\begin {align*}
\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{j+1}\end{array}\right)\ \mbox {with}\\
L_{j+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
\end{align*}
\textcolor{blue}{This step is performed by function \texttt{addCollisionConstraint}}.

@@

Thus the linearized constraint Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ 
is built by matrix 
blocks using the Jacobian $\frac{\partial f}{\partial \conf}$ expressed in 
each of the two waypoints and $\beta$. This stage \textcolor{blue}{is performed by \texttt{computeCollisionConstraint}}.

\subsection{Convergence Analysis and algorithm refinment}

\begin {figure}
  \centerline {
    \def\svgwidth {.7\linewidth}
    \graphicspath{{./images/}}
    \input {images/convergence.pdf_tex}
  }
  \caption {\textcolor{red}{Illustration of linearly constrained quadratic program on a problem analogous to the path shortening problem: we want to minimize the quadratic cost $\frac{1}{2}\|\xx-\xx^*\|^2, \xx\in\real^2$ under the non-linear constraint $f(\xx)\leq 0$. $\alpha_{init}=\frac{1}{4}$. The algorithm starts from $\xx_0$. The first iterate is $\xx_1$ which satisfies the constraint. The second iterate is $\xx_2$ that does not satisfy the constraint. The algorithm backtracks to $\xx_1$ and inserts linear constraint $L_1: \frac{\partial f}{\partial \xx} (\xx_1)(\xx - \xx_1) = 0$. $\xx_3$ is the global minimum under $L_1$. As $\xx_3$ does not satisfy the constraint, the algorithm moves to $\xx_4$ that satisfies the constraints, and then to $\xx_5$ that does not. The algorithms backtracks to $\xx_4$, inserts constraint $L_2: \frac{\partial f}{\partial \xx} (\xx_2)(\xx - \xx_2) = 0$, and returns $\xx_4$ as a solution since the dimension of the search space is 0. Notice that the same non-linear constraint may give rise to several linear constraints. The convergence of the algorithm relies on the fact that the kernel of constraint $L2$ is not contained in the kernel of the current constraints ($L_1$ only here). The convergence analysis can be roughly summarized as follows. $L2$ to be linearly dependent of $L_1$ requires that $f$ is stationary along $L_1$ at $\xx_4$. This is unlikely (but possible) since $\xx_4$ is not far from the boundary of the domain defined by $f(\xx) \leq 0$. If $L_2$ was not linearly independent from $L_1$, The algorithm would keep searching new iterates between $\xx_5$ and $\xx_4$. If by any chance constraint $f$ linearized around each of those collision-free iterates was each time linearly dependent from $L_1$ the iterates would converge to the boundary of the domain defined by $f(\xx) \leq 0$. By continuous differentiability of $f$, this would mean that $f$ is stationary at the boundary. In other words, the straight line passing by $\xx_1$ and $\xx_3$ would cross the boundary tangently (as in bottom right picture). This is possible but unlikely, unless the problem has been defined as such on purpose.}}
  \label {fig:convergence}
\end{figure}

@@

From the definition of $\Jf_{i}$, it is straightforward that:
\begin{equation}\label{eq:decreasing-kernel}
\kernel \Jf_{i+1} \subset \kernel \Jf_{i}
\end{equation}
Let us assume that:
\begin{equation}\label{eq:assumption-1}
L_{j+1}\p_i \not= 0
\end{equation}
We will elaborate later on this assumption. This means that 
$\p_{i}\notin\kernel \Jf_{i+1}$, and as $\p_{i}\in\kernel \Jf_{i}$, then:
$$
\kernel \Jf_i \not= \kernel \Jf_{i+1}
$$
From~(\ref{eq:decreasing-kernel}), we deduce that:
$$
\dim (\kernel \Jf_{i+1}) < \dim (\kernel \Jf_i)
$$
This result proves that under assumption~(\ref{eq:assumption-1}), each additional constraint is linearly independent from the 
previous ones, and thus our algorithm terminates in a finite number of iterations.

@@

\subsubsection{Note about assumption~(\ref{eq:assumption-1})}%
\noindent
\textcolor{red}{As in the previous section, we assume that $\xx_i$ is collision-free and that $\xx_{i+1}$ is in collision at abscissa $\tcolli$.
According to~(\ref{eq:f}), the evaluation of the constraint function $f$ along the iteration line $\xx_i + t\alpha_i \p_i,\ t\in [0,1]$ going from $\xx_i$ to $\xx_{i+1}$ can be written as follows:
\begin{equation}\label{eq:f-along-pi}
f (\xx_i + t\alpha_i \p_i) = g \left((\xx_i + t\alpha_i \p_i)(\tcolli)\right).
\end{equation}
The argument of function $g$ above is a trajectory in the robot configuration space that we denote by $\traj$:
\begin{equation}\label{eq:trajectory}
\traj (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (\tcolli), \ t \in [0,1]
\end{equation}
The trajectory $\traj(t)$ is defined by taking the constant abscissa $\tcolli$ and by 
moving from 
path $\xx_i$ along step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
Substituting~(\ref{eq:trajectory}) into~(\ref{eq:f-along-pi}) and differentiating with respect  to $t$ yields
\begin{align}
f (\xx_i + t\alpha_i \p_i) &= g (\traj (t)) \\
\label{eq:df/dx}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i + t\alpha_i \p_i)\p_i &=
\frac{d}{dt}g (\traj (t))
\end{align}
From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $P_2$ in reference frame $\body_1$ projected on vector $\U$.
}

\textcolor{red}{
Therefore the following expressions
\begin{align*}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i)\p_i = \frac{d}{dt}g (\traj (0)) & &
\alpha_i \frac{\partial f}{\partial \xx} (\xx_{i+1})\p_i =\frac{d}{dt}g (\traj (1))
\end{align*}
correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $P_2$ in reference frame $\body_1$. 
Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0. According to the geometric interpretation, Assumption~(\ref{eq:assumption-1}) is violated if and only if $\velocity_{\po_2/\body_1}$ is orthogonal to $\U$. Although very unlikely, this case might appear for some new constraint. In this case, inserting constraint $L_{j+1}$ is useless since
$$
\kernel \Jf_i = \kernel \Jf_{i+1}.
$$
}

\subsubsection {Algorithm refinment}

\textcolor{red}{
When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop in order to find a new constraint that is linearly independent. This loop is performed by Algorithm~\ref{algo:solveRedundant} and illustrated in Figure~\ref{fig:convergence-diagram}. The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy. A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop. New iterations are chosen in the middle of this pair. 
\begin{itemize}
\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
\end{itemize}
In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
\begin{enumerate}
\item at some point in the loop the new constraint is linearly independent from the previous constraints. The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
\end{enumerate}
In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$. $\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$. Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
}

\begin{figure}
	\centering
	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
	\caption{\textcolor{blue}{Considering a trajectory 
	$\conf (t) = (\xx_{Free} + t\, \alpha \p)(\kappa)$ 
	with $\xx_{Coll} = \xx_{Free} + \alpha\p$ ($\xx_{Free}$, $\xx_{Coll}$, $\p$ 
	and $\kappa$ respectively collision-free path, in-collision path, iteration step 
	and collision-abcissa).
	The trajectory is represented in the configuration space (left) and in the 
	workspace with the colliding bodies (right).}
	Note that a case with 
	$(\velocity_{\po_2/\body_1}(\xx_{Free})|\mathbf{u})=0$ is considered, so $\p$ already complies 
	with the constraint that should be added if no redundancy was tracked. Therefore, 
	Algorithm~\ref{algo:solveRedundant} will produce one intermediate step
	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
	$(\velocity_{\po_2/\body_1}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
	returning the new constraint.}
	\label{fig:convergence-diagram}
\end{figure}
